---
title: "PA1_templeta"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*PROJECT 1: REPRODUCIBLE RESEARCH

This code loads the data, processes, cleans the data and gets a summary of the data

```{r simulation, echo =TRUE}
library("dplyr")
setwd("C:/Users/user/Desktop/project1")
data = read.csv("activity.csv", header = TRUE)
activity_data<-na.omit(data)
summary(activity_data)
```

Total number of steps taken per day 
```{r}
grp_data = activity_data %>% group_by(date)  %>%
   summarise(total_steps_per_day = sum(steps),
             .groups = 'drop')
grp_data
```
Histogram of the total number of steps taken each day
```{r}
hist(grp_data$total_steps_per_day, col = "blue")
```

Mean and median number of steps taken each day
```{r}
mean_steps<-mean(grp_data$total_steps_per_day)
median_steps<-median(grp_data$total_steps_per_day)
mean_steps
median_steps
```

Average daily pattern.. Time series plot of the average number of steps taken
```{r}
steps_per_interval<-aggregate(steps~interval, data=activity_data, mean)
plot(steps~interval, data=steps_per_interval, type="l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
interval_with_max_step<-steps_per_interval[which.max(steps_per_interval$steps), ]$interval
interval_with_max_step
```

Calculate total number of missing values
```{r}
total_no_missing_values <- sum(is.na(data$steps))
total_no_missing_values
```

Code to describe and show a strategy for imputing missing data
```{r}
library(mice)
library(Hmisc)
data$imputed_steps <- with(data, impute(data$steps, mean))
md.pattern(data)

grp_imputed_data = data %>% group_by(date)  %>%
   summarise(imputed_steps_per_day = sum(imputed_steps),
             .groups = 'drop')
hist(grp_imputed_data$imputed_steps_per_day, col = "blue")
```
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
data$date <- as.Date(strptime(data$date, format="%Y-%m-%d"))
data$day <- weekdays(data$date)
for (i in 1:nrow(data)) {
    if (data[i,]$day %in% c("Saturday","Sunday")) {
        data[i,]$day<-"weekend"
    }
    else{
        data[i,]$day<-"weekday"
    }
}
stepsByDay <- aggregate(data$imputed_steps ~ data$interval + data$day, data, mean)
```
Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```